---
title: 'This is a note for working on a GEO data'
author: 'Jianying Li'
date: '30 January 2020'
output:
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: readable
    highlight: tango
---

# Introduction

Yicheng helped David in Dr. Chen's lab to write a SAM-GSA code for microarray analysis. Here is a note for such analysis.

Here are the files that are needed for the analysis

* Normalized microarray data
* GSA geneset
* This program.

# Load and check data

```{r, message = FALSE}
# Load packages
library(samr)
library(GSA)
library(Biobase)
library(GEOquery)
```

Now that our packages are loaded, let's read in and take a peek at the data.
Download a sample Microarray Data from GEO, with code inherited from GEO2R
```{r, message=FALSE, warning=FALSE}


# load series and platform data from GEO

gset <- getGEO("GSE18842", GSEMatrix =TRUE, getGPL=FALSE)
if (length(gset) > 1) idx <- grep("GPL570", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]

# group names for all samples in a series
gsms <- paste0("10101101010101011010101010101010010110101010101101",
               "01000001010101010101010010110101110101010")
sml <- c()
for (i in 1:nchar(gsms)) { sml[i] <- substr(gsms,i,i) }
sml <- paste("G", sml, sep="")  #set group names

# order samples by group
ex <- exprs(gset)[ , order(sml)]
sml <- sml[order(sml)]
fl <- as.factor(sml)
labels <- c("control","tumor")

```

# Examine by plot
```{r, message=FALSE, warning=FALSE}
# set parameters and draw the plot
palette(c("#dfeaf4","#f4dfdf", "#AABBCC"))
dev.new(width=4+dim(gset)[[2]]/5, height=6)
par(mar=c(2+round(max(nchar(sampleNames(gset)))/2),4,2,1))
title <- paste ("GSE18842", '/', annotation(gset), " selected samples", sep ='')
boxplot(ex, boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=fl)
legend("topleft", labels, fill=palette(), bty="n")
```



# Running SAM analysis

```{r, message=FALSE, warning=FALSE}
#Getting the class label

y <- rep ("1", dim(ex)[2])
y[sml %in% "G1"] = 2
y
samfit <- SAM(as.matrix(ex), y, resp.type = "Two class unpaired")
plot(samfit)
```



Prepare files for GSA now

```{r, message=FALSE, warning=FALSE}

affyProb2Symbols <- function (aListOfProbes, db = hgu133plus2.db)
{
  require(magrittr)
  require(hgu133plus2.db)
  probeAnnoted <- AnnotationDbi::select(
    x       = db,
    keys    = aListOfProbes,
    columns = c("PROBEID", "ENSEMBL", "ENTREZID", "SYMBOL"),
    keytype = "PROBEID"
  )
  return (probeAnnoted )
}

#x = as.matrix(ex)

geneAnnoated <- affyProb2Symbols(tolower(rownames(ex)))
library(tidyverse)
filtered = geneAnnoated %>% distinct(PROBEID, .keep_all = TRUE)
#genenames = filtered$SYMBOL
#paste ("g", 1:dim(ex)[1], sep="")

data <- list(x = as.matrix(ex), y = y, geneid = paste ("g", 1:dim(ex)[1], sep=""), genenames = filtered$SYMBOL, logged2 = TRUE)

```

# Working on GSA
## Basic set up

```{r , message=FALSE, warning=FALSE}

## Setting the random seed for future validation
set.seed(1234567)

```


## Load gene set database


```{r , message=FALSE, warning=FALSE}


clean.gmt.data <- function (db)
{
# db <- c1 
# str(db)
# db$geneset.names
# db$geneset.descriptions
 temp.geneset <- c(list())
 for (i in 1:length(db$genesets))
 {
   temp.geneset[[i]] <- db$genesets[[i]]
   if (length(which(temp.geneset[[i]]=="")) > 0 )
   {
      temp.geneset[[i]]  <- temp.geneset[[i]][-which(temp.geneset[[i]]=="")]
   }
 }
 db$genesets <- temp.geneset 
 return (db)
}

## Rename "Knowledge based gene sets (20161005).gmt" to geneset_2.gmt
#c1 <- GSA.read.gmt("genset_2.gmt") 

# original file
c1 <- GSA.read.gmt("X:/myGit/shortcourse/GSA_R/Knowledge_based_gene_sets_20161005.gmt")
#c1 <- GSA.read.gmt("cleaningData/genset_2_correct.txt")
#c1 <- GSA.read.gmt("cleaningData/genset_2_correct_from_1.txt")
#c1 <- GSA.read.gmt("cleaningData/genset_2.txt")
# GSA.read.gmt constructs a list of 3

c1.cleaned <- clean.gmt.data(c1)
str(c1)
str(c1$genesets)

print ("Gene sets in database file (20161005).gmt")
c1$geneset.names

```

## Run GSA


```{r  results = 'hide', message=FALSE, warning=FALSE}
c1 <- c1.cleaned


#data <- list(x = as.matrix(ex), y = y, geneid = paste ("g", 1:dim(ex)[1], sep=""), genenames = filtered$SYMBOL, logged2 = TRUE)


gsafit <- GSA(as.matrix(ex), y, c1$genesets, genenames = filtered$SYMBOL, resp.type = "Two class unpaired", nperms = 1000)

## Do I need the unique gene symbol??
#gsafit <- GSA(x, y, c1$genesets, unique(z$GeneSymbol), resp.type = "Two class unpaired", nperms = 1000)


str(gsafit)
plot(gsafit$GSA.scores)

```



```{r ,  message=FALSE, warning=FALSE}
gsafit$fdr.lo
gsafit$fdr.hi

# To many and will NOT print it to the html file
# gsafit$gene.scores

GSA.plot(gsafit)

GSA.listsets(gsafit, c1$geneset.names, FDRcut=.5)

GSA.correlate(c1$genesets, filtered$SYMBOL)
gsa2 <- GSA.func(as.matrix(ex), y, c1$genesets, filtered$SYMBOL, resp.type = "Two class unpaired")
#gsa2 <- GSA.func(x, y, c1$genesets, unique(z$GeneSymbol), resp.type = "Two class unpaired")
str(gsa2)



```

## More on GSA


```{r , message=FALSE, warning=FALSE}



#for loop from 1:length(gsals$negative[,1])
GSA.genescores(5, c1$genesets, gsafit, z$GeneSymbol)
GSA.plot(gsafit, fac = 1, FDRcut = 1)

gsals <- GSA.listsets(gsafit, c1$geneset.names, FDRcut=.5)
str(gsals)
str(gsals$negative)
gsals$negative[,1]
length(gsals$negative[,1])
as.numeric(gsals$negative[,1][1])
GSA.genescores(as.numeric(gsals$negative[,1][1]), c1$genesets, gsafit, z$GeneSymbol)


```


## Export results


```{r , message=FALSE, warning=FALSE}
## sink() function does not work well with rmarkdown, 

#sink("V1-GSA-TF-negative.txt")

for (i in 1:length(gsals$negative[,1])){
  print(gsals$negative[,2][i])
  print(GSA.genescores(as.numeric(gsals$negative[,1][i]),
                 c1$genesets, gsafit, filtered$SYMBOL))
  cat("\n")
  
}
#sink()

#sink("V1-GSA-TF-positive.txt")

for (i in 1:length(gsals$positive[,1])){
  print(gsals$positive[,2][i])
  print(GSA.genescores(as.numeric(gsals$positive[,1][i]),
                       c1$genesets, gsafit, filtered$SYMBOL))
  cat("\n")
  
}
#sink()

#print GSA -TF output
#sink("V1-GSA-TF-output.txt")
print(GSA.listsets(gsafit, c1$geneset.names, FDRcut=.5))
#sink()

#Print negative names

#sink("v2-negativenames.txt")
print ("Negaivie names")
for (i in as.numeric(gsals$negative[,1])){
  print(c1$geneset.names[i])
  
}
#sink()

#print positivenames

#sink("V1-positivenames.txt")
print ("Positivie names")
for (i in as.numeric(gsals$positive[,1])){
  print(c1$geneset.names[i])
  
}
#sink()

```


# Conclusion

This is a valid way to document analysis.

---


